---
title: "Introduction to cDiD with Multiple Time Periods"
author: "Joel Cuerrier"
date:  "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to cDiD with Multiple Time Periods}
  %\VignetteEngine{knitr::rmarkdown}    
  %\VignetteEncoding{UTF-8}
---

****
```{r}
# browseVignettes("did")
```

<!-- rmarkdown::pandoc_version() -->



```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Introduction 
This vignette provides an introductory overview of utilizing Chained-Difference-in-Differences (cDiD) designs to identify and estimate the average impact of engaging in a treatment, with a specific emphasis on leveraging functionalities from the cdid package. 

You can find the associated background article for further reference here.

* The cdid package allows for multiple periods, variation in treatment timing, treatment effect heterogeneity, general missing data patterns, and sample selection on observables.
* The chained DiD rests on a parallel trends assumption, conditioned on being sampled. 
* The chained DiD is robust to some forms of attrition caused by unobservable heterogeneity.
* We identify three main advantages of our approach: (1) it does not require having a balanced panel subsample as it is the case with a standard DiD; (2) identifying assumption allows for sample selection on time-persistent unobservable (or latent) factors, unlike the cross-section DiD which treats the sample as repeated cross-sectional data; and (3) it may also deliver efficiency gains compared to existing methods, notably when the outcome variable is highly time-persistent.
* The cdid package is designed to provide group-time average treatment effects. Future updates to the package will expand its capabilities to include event-study type estimates, which encompass treatment effect parameters corresponding to various durations of exposure to the treatment, as well as overall treatment effect estimates.

## Examples with numerical simulations

We propose a simulation design adapted from the first section. Let us specify the
potential outcome as a components of variance:


$$
Y_{it}(D_i) = \alpha_i + \delta_t + \sum_{\tau=2}^{t} \beta_{\tau} D_{i\tau} + \varepsilon_{it}
$$

Where $D_{i\tau} ∈ {0, 1}$ denotes whether individual i has been treated in $\tau$ or earlier. Let us assume that $\tau ∈ {0, ..., T + 1}$ and treatments can only occur in t ≥ 2 so that G ∈ {2, ..., T + 1}. The data generating process is characterized by the following assumptions:

• The individual-specific unobservable heterogeneity is iid gaussian: $\alpha_i ∼ N(1, \alpha^2_{\alpha})$, where $\alpha^2 = 2;$
• The time-specific unobservable heterogeneity is iid gaussian: $\delta_t ∼ N(1, 1)$;
• The treatment effect is constant over time: $\beta_{\tau} = 1$;
• The error term is iid gaussian: $\varepsilon_{it} ∼ N(0, \alpha^2_{\varepsilon})$;
• The probability to receive the treatment at time g, conditional on being treated at g or in the control group, is defined as: 
$$Pr((G_{ig} = 1|X_i, \alpha_i, G_{ig} + C_i = 1)) = \frac{1}{1 + \exp(\theta_0 + \theta_1 X_i + \theta_2 \alpha_i \times g)}$$.
Where $X_i ∼ N(1, 1)$ is observable for every i, unlike $\alpha_i$, and $\theta_0 = −1, \theta_1 = 0.4$ and $\theta_2 = 0$ or $\theta_2 = 0.2$. In the latter case, the treatment probability varies with treatment timing and the unobserved individual heterogeneity;
• The sampling probability in the consecutive periods t, t + 1 conditional on $\alpha_i$ is given by:
$$Pr((S_{it} + 1 = 1|\alpha_i)) = \frac{1}{1 + \exp(\lambda_0 + \lambda_1 \alpha_i \times t)}$$
With $\lambda_0 = −1$, and $\lambda_1 = 0$ or $\lambda_1 = 0.2$, so that the sampling process can also vary with time and the unobserved individual heterogeneity.

We simulate the sampled data in two steps. First, we generate a population sample
for each period t to represent individuals that are either treated at t or in the control
group. Second, we sample from this population using the specified process. We

1. Generate a population of individuals
(a) Draw $N = 2 x max $\frac{n}{E_{\alpha}[Pr(S_{itt+1})]}$ individuals per period in order to have $T + 2$ population samples of $ N $ individuals, where each individual is characterized by a vector $(\alpha_i, \delta_t, X_i, \epsilon_{it})$.
(b) Separately for each population sample g, draw a uniform random number $(\xi_i \in [0, 1])$ per individual. If $\xi_i \leq Pr(G_{it} = 1|X_i, \alpha_i, G_{it} + C_i = 1)$, then set $(G_{ig} = 1, C_i = 0)$, otherwise set $(G_{ig} = 0, C_i = 1)$.
(c) Compute $Y_{it}$ from $(\alpha_{i}, \delta_{t}, X_{i}, ε_{it}, G_{i0}, ..., _{iT+1}, C_{i})$;

2. Sample from this population
(a) Draw a uniform random number $η_{it} ∈ [0, 1]$ per individual i and period t. If $η_{it} ≤ Pr(S_{itt+1} = 1|\alpha_{i}), then set S_{itt+1} = 1$ and $S_{iττ+1} = 0$ for $τ /neq t;
(b) Draw (without replacement) n individuals per period t from the population for which $S_{itt+1} = 1$;
(c) Compute the different estimators.
(d) Repeat steps 1(b)-2(c) 1,000 times and report the mean and standard deviation of the estimators.

## Estimating the average treatment effect 
# Building the dataset

```{r}
set.seed(123)
#importing the package
library(cdid)
#import library from githuib
devtools::install_github("joelcuerrier/cdid")
# Using the function `fonction_simu_attrition` from the `cdid` package, we simulate the data for the following parameter values: $\theta_2 = 0.2, \lambda_1 = 0.2, \sigma_{\alpha} = 2, \sigma_{\epsilon} = 0.5$.


data=data_sim=fonction_simu_attrition(nbsimu = 1, theta2_alpha_Gg=0.2, lambda1_alpha_St=0.2, sigma_alpha=2, sigma_epsilon=0.5)
head(data)
```
