---
title: "Introduction to cDiD with Multiple Time Periods"
author: "Joel Cuerrier"
date:  "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to cDiD with Multiple Time Periods}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
options(repos = c(CRAN = "https://cran.rstudio.com"))
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
install.packages("pandoc")
library(pandoc)
```
<!-- rmarkdown::pandoc_version() -->


# Attribution of Vignettes to Original Author
The vignettes included in this library are based on those authored by [Brantly Callaway and Pedro H.C. Sant'Anna] for the 'did' package. We have adapted the content for use in this library to provide users with accessible and informative documentation. The original vignettes authored by [Brantly Callaway and Pedro H.C. Sant'Anna] can be found in the 'did' package documentation [https://bcallaway11.github.io/did/articles/did-basics.html].

# Introduction 
****
* This vignette provides an introductory overview of utilizing Chained-Difference-in-Differences (cDiD) designs to identify and estimate the average impact of engaging in a treatment, with a specific emphasis on leveraging functionalities from the cdid package. 

* The cdid package allows for multiple periods, variation in treatment timing, treatment effect heterogeneity, general missing data patterns, and sample selection on observables.

* The chained DiD rests on a parallel trends assumption, conditioned on being sampled. 

* The chained DiD is robust to some forms of attrition caused by unobservable heterogeneity.

* We identify three main advantages of our approach: (1) it does not require having a balanced panel subsample as it is the case with a standard DiD; (2) identifying assumption allows for sample selection on time-persistent unobservable (or latent) factors, unlike the cross-section DiD which treats the sample as repeated cross-sectional data; and (3) it may also deliver efficiency gains compared to existing methods, notably when the outcome variable is highly time-persistent.

* The cdid package is designed to provide group-time average treatment effects. Future updates to the package will expand its capabilities to include event-study type estimates, which encompass treatment effect parameters corresponding to various durations of exposure to the treatment, as well as overall treatment effect estimates.

# Examples with numerical simulations
****
We propose a simulation design adapted from the first section. Let us specify the
potential outcome as a components of variance:

$$
Y_{it}(D_i) = \alpha_i + \delta_t + \sum_{\tau=2}^{t} \beta_{\tau} D_{i\tau} + \varepsilon_{it}
$$

Where $D_{i\tau} ∈ {0, 1}$ denotes whether individual i has been treated in $\tau$ or earlier. Let us assume that $\tau ∈ {0, ..., T + 1}$ and treatments can only occur in t ≥ 2 so that G ∈ {2, ..., T + 1}. The data generating process is characterized by the following assumptions:

* The individual-specific unobservable heterogeneity is iid gaussian: $\alpha_i ∼ N(1, \alpha^2_{\alpha})$, where $\alpha^2 = 2;$
* The time-specific unobservable heterogeneity is iid gaussian: $\delta_t ∼ N(1, 1)$;
* The treatment effect is constant over time: $\beta_{\tau} = 1$;
* The error term is iid gaussian: $\varepsilon_{it} ∼ N(0, \alpha^2_{\varepsilon})$;
* The probability to receive the treatment at time g, conditional on being treated at g or in the control group, is defined as: 
$$Pr(G_{ig} = 1|X_i, \alpha_i, G_{ig} + C_i = 1) = \frac{1}{1 + \exp(\theta_0 + \theta_1 X_i + \theta_2 \alpha_i \times g)}$$.
Where $X_i ∼ N(1, 1)$ is observable for every i, unlike $\alpha_i$, and $\theta_0 = −1, \theta_1 = 0.4$ and $\theta_2 = 0$ or $\theta_2 = 0.2$. In the latter case, the treatment probability varies with treatment timing and the unobserved individual heterogeneity;
* The sampling probability in the consecutive periods t, t + 1 conditional on $\alpha_i$ is given by:
$$Pr((S_{it} + 1 = 1|\alpha_i)) = \frac{1}{1 + \exp(\lambda_0 + \lambda_1 \alpha_i \times t)}$$
With $\lambda_0 = −1$, and $\lambda_1 = 0$ or $\lambda_1 = 0.2$, so that the sampling process can also vary with time and the unobserved individual heterogeneity.

We simulate the sampled data in two steps. First, we generate a population sample
for each period t to represent individuals that are either treated at t or in the control
group. Second, we sample from this population using the specified process. 

1. Generate a population of individuals
(a) Draw $N = 2 x max $\frac{n}{E_{\alpha}[Pr(S_{itt+1})]}$ individuals per period in order to have $T + 2$ population samples of N individuals, where each individual is characterized by a vector $(\alpha_i, \delta_t, X_i, \epsilon_{it})$.
(b) Separately for each population sample g, draw a uniform random number $(\xi_i \in [0, 1])$ per individual. If $\xi_i \leq Pr(G_{it} = 1|X_i, \alpha_i, G_{it} + C_i = 1)$, then set $(G_{ig} = 1, C_i = 0)$, otherwise set $(G_{ig} = 0, C_i = 1)$.
(c) Compute $Y_{it}$ from $(\alpha_{i}, \delta_{t}, X_{i}, ε_{it}, G_{i0}, ..., _{iT+1}, C_{i})$;

2. Sample from this population
(a) Draw a uniform random number $η_{it} ∈ [0, 1]$ per individual i and period t. If $η_{it} ≤ Pr(S_{itt+1} = 1|\alpha_{i}), then set S_{itt+1} = 1$ and $S_{iττ+1} = 0$ for $τ /neq t$;

(b) Draw (without replacement) n individuals per period t from the population for which $S_{itt+1} = 1$;
(c) Compute the different estimators.
(d) Repeat steps 1(b)-2(c) 1,000 times and report the mean and standard deviation of the estimators.

```{r}
# Load the cdid package
remotes::install_github("joelcuerrier/cdid", ref = "main", build_vignettes = FALSE, force = TRUE)
library(cdid)
?cdid
?cdid::chained
?cdid::GMM

# Importing the data
data0 = fonction_simu_attrition()

# Display the data
head(data0)
```

```{r}
library(did)
did.results = did:: att_gt(
  yname="Y", 
  tname="date",
  idname = "id",
  gname = "date_G",
  xformla = ~X,
  data = data0,
  weightsname = NULL,
  allow_unbalanced_panel = TRUE,
  alp = 0.05,
  bstrap = TRUE,
  cband = TRUE,
  biters = 1000,
  clustervars = NULL,
  est_method = "dr",
  base_period = "varying",
  print_details = FALSE,
  pl = FALSE,
  cores = 1
)

#Dynamic Effect of the treatment
agg.es.did = aggte(MP = did.results, type = 'dynamic')

# 3. Print results
agg.es.did
```


```{r}
# # Load the cdid package


# # 0. We generate data.

# source("R/fonction_simu_attrition.R")
# # 1. Chained DiD
# chained.results = chained(
#                     data=data0,      
#                     yname="Y",
#                     tname="date",
#                     idname="id",
#                     gname="date_G",
#                     control_group=c("notyettreated"), #either "nevertreated" or "notyettreated"
#                     treated='treated',
#                     xformla=~X, 
#                     weightsname="S",  
#                     bstrap=FALSE, 
#                     biters=1000,
#                     cband=TRUE)
# # See the ATT g,t
# chained.results

# # 2. Aggregation
# agg.es.chained <- aggte(MP = chained.results, type = 'dynamic')

# # 3. Print results
# agg.es.chained
```


```{r}

# 1. GMM DiD
gmm.results = GMM(
                    data=data0,      
                    yname="Y",
                    tname="date",
                    idname="id",
                    gname="date_G",
                    control_group=c("notyettreated"), #either "nevertreated" or "notyettreated"
                    treated='treated',
                    xformla=~X, 
                    weightsname="S",  
                    bstrap=FALSE, 
                    biters=1000,
                    cband=TRUE)

# See the ATT g,t
gmm.results

# 2. Aggregation
agg.es.gmm <- aggte(MP = gmm.results, type = 'dynamic')

# 3. Print results
agg.es.gmm
```


```{r}
# # Combine group and time results
# did.attgt <- data.frame(group = did.results$group, t = did.results$t, did.att = did.results$att)
# gmm.attgt <- data.frame(group = gmm.results$group, t = gmm.results$t, gmm.att = gmm.results$att)
# did.se <- data.frame(group = did.results$group, t = did.results$t, did.se = did.results$se)
# gmm.se <- data.frame(group = gmm.results$group, t = gmm.results$t, gmm.se = gmm.results$se)

# # Merge all results
# results.attgt <- Reduce(function(x, y) merge(x, y, by = c("group", "t"), all = FALSE),
#                         list(did.attgt, gmm.attgt, did.se, gmm.se))


# # Aggregate results
# agg.es.did <- aggte(MP = did.results, type = 'dynamic')
# agg.es.gmm <- aggte(MP = gmm.results, type = 'dynamic')

# did.agg.attgt <- data.frame(egt = agg.es.did$egt, did.att = agg.es.did$att, did.se = agg.es.did$se)
# gmm.agg.attgt <- data.frame(egt = agg.es.gmm$egt, gmm.att = agg.es.gmm$att, gmm.se = agg.es.gmm$se)

# # Merge aggregated results
# results.agg.attgt <- merge(did.agg.attgt, gmm.agg.attgt, by = "egt", all = FALSE)

# # Reorder columns in results.agg.attgt to match results.attgt
# results.agg.attgt <- results.agg.attgt[, c("egt", "did.att", "gmm.att", "did.se", "gmm.se")]

# # Display results
# results.attgt
# results.agg.attgt
```


```{r}
# Combine group and time results
did.attgt <- data.frame(group = did.results$group, t = did.results$t, did.att = did.results$att)
gmm.attgt <- data.frame(group = gmm.results$group, t = gmm.results$t, gmm.att = gmm.results$att)

did.se <- data.frame(group = did.results$group, t = did.results$t, did.se = did.results$se)
gmm.se <- data.frame(group = gmm.results$group, t = gmm.results$t, gmm.se = gmm.results$se)

did.agg.attgt <- data.frame(egt = agg.es.did$egt, did.att = agg.es.did$att, did.se = agg.es.did$se)
gmm.agg.attgt <- data.frame(egt = agg.es.gmm$egt, gmm.att = agg.es.gmm$att, gmm.se = agg.es.gmm$se)





```

```{r}

# compare.results <- function(...){
#   # List of data frames passed as arguments
#   results_list <- list(...)
  
#   # Check if each result has the required columns: "group" and "t"
#   for (df in results_list) {
#     if (!all(c("group", "t") %in% colnames(df))) {
#       stop("Each result data frame must contain 'group' and 't' columns")
#     }
#   }
  
#   # Merge all results by 'group' and 't'
#   combined_results <- Reduce(function(x, y) merge(x, y, by = c("group", "t"), all = FALSE), results_list)
  
#   return(combined_results)
# }

# compare.results(did.attgt, gmm.attgt, did.se, gmm.se)

```

```{r}

# compare.agg.results <- function (...){
# results_list <- list(...)
  
#   # Check if each result has the required columns: "group" and "t"
#   for (df in results_list) {
#     if (!all(c("egt") %in% colnames(df))) {
#       stop("Each result data frame must contain 'group' and 't' columns")
#     }
#   }
# # Merge all results by 'group' and 't'
#   combined_results <- Reduce(function(x, y) merge(x, y, by = c("egt"), all = FALSE), results_list)
  
#   return(combined_results)
# }

# compare.agg.results(did.agg.attgt, gmm.agg.attgt)

```

```{r}

compare.results <- function(..., agg = FALSE) {
  # List of data frames passed as arguments
  results_list <- list(...)
  
  # Check if each result has the required columns
  for (df in results_list) {
    if (agg) {
      # If agg is TRUE, check only for the "egt" column
      if (!"egt" %in% colnames(df)) {
        stop("Each result data frame must contain 'egt' column when agg = TRUE")
      }
    } else {
      # If agg is FALSE, check for both "group" and "t" columns
      if (!all(c("group", "t") %in% colnames(df))) {
        stop("Each result data frame must contain 'group' and 't' columns when agg = FALSE")
      }
    }
  }
  
  # Merge all results by 'egt' (for aggregated data) or 'group' and 't' (for regular data)
  if (agg) {
    # Merge based on "egt"
    combined_results <- Reduce(function(x, y) merge(x, y, by = "egt", all = FALSE), results_list)
  } else {
    # Merge based on "group" and "t"
    combined_results <- Reduce(function(x, y) merge(x, y, by = c("group", "t"), all = FALSE), results_list)
  }
  
  # Return the merged results
  return(combined_results)
}

# Example usage for aggregated data
results_agg <- compare.results(did.agg.attgt, gmm.agg.attgt, agg = TRUE)
print(results_agg)

# Example usage for non-aggregated data
results_non_agg <- compare.results(did.attgt, gmm.attgt, agg = FALSE)
print(results_non_agg)



```


<!-- import(dplyr)
import(BMisc)
importFrom(BMisc, rhs.vars)
import(tidyr)
importFrom(tidyr, pivot_wider)
import(Matrix)
importFrom(Matrix, sparseMatrix)
import(openxlsx)
import(data.table)
import(jsonlite)
import(tidyr)
import(did) -->